---
title: "Aides Drug"
author: "April Parry"
date: "Summer 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

```{r}
Charlie <- filter(triUmp, DrugUser == "yes")
```


```{r eval=FALSE}
Charlie<-Charlie %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

I need to use the filter command to get a subset of data of the people in triump down to smaller set of data comprised of DrugUser only.


<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!--


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data = Charlie)
barchartGC(~infected + treatment,data=Charlie, type="percent")
```

These two plots explain how people got the disease aids data by frequency and percentage.

```{r}
library(ggplot2)

# Assuming the dataframe is named "Charlie"
# Replace "infected" and "treatment" with the actual column names in your dataframe

# Convert the factor variables to factors if they are not already
Charlie$infected <- as.factor(Charlie$infected)
Charlie$treatment <- as.factor(Charlie$treatment)

# Plot the barcharts
ggplot(Charlie, aes(x = infected, fill = treatment)) +
  geom_bar(position = "dodge", color = "black") +
  labs(x = "Infected", y = "Count") +
  ggtitle("Bar Chart of infected and Treatment") +
  scale_fill_manual(values = c("drug" = "blue", "placebo" = "red")) +
  theme(legend.position = "bottom")

ggplot(Charlie, aes(x = treatment, fill = treatment)) +
  geom_bar(position = "dodge", color = "black") +
  labs(x = "Infected", y = "Count") +
  ggtitle("Bar Chart of Treatment") +
  scale_fill_manual(values = c("drug" = "blue", "placebo" = "red")) +
  theme(legend.position = "bottom")

```

The first plot shows how many people got infected.  It shows that the drug is working against AIDS. The relationship is 3 quarters of the people have got the placebo. The ratio is 50/50 so the drug is preventing those who took it from getting AIDS.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
table2 <- xtabs(~infected + treatment, data=Charlie)
rowPerc(table2)
colPerc(table2)
```

Of the people that go aids 24.29 took the drug .75.71% took the placebo. From the people that haven't got anything was 47.59% to 52.41% This is is a sign of a drug that looks like it is working. 
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table2)
chisqtestGC(table2)
```

The probability of getting this result is an extremely low number of .0000000423. This number leads us to reject the null hypothesis and tells us the drug is working. The small p value shows something is happening.



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

In this write up we used my tricks rmd.Dplyr to make the triUmp group into an even smaller group called DRUGUSR. The drug users had a positive results showing us that the drug was in fact effective, and we can see those results in the graphical and numerical data.The results from the inferential test confirms that the drug did in fact have a positive result and was a working drug.
