---
title: "Aides Drug"
author: "April Parry"
date: "Summer 2023"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

```{r}
Charlie <- filter(triUmp, DrugUser == "yes")
```


```{r eval=FALSE}
Charlie<-Charlie %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

I need to use the filter command to get a subset of data of the people in triump down to smaller set of data comprised of DrugUser only.

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!--



# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data = Charlie)
barchartGC(~infected + treatment,data=Charlie, type="percent")
```

These two plots explain how people got the disease aids data by frequency and percentage.

```{r}
library(ggplot2)

# Assuming the dataframe is named "Charlie"
# Replace "infected" and "treatment" with the actual column names in your dataframe

# Convert the factor variables to factors if they are not already
Charlie$infected <- as.factor(Charlie$infected)
Charlie$treatment <- as.factor(Charlie$treatment)

# Plot the barcharts
ggplot(Charlie, aes(x = infected, fill = treatment)) +
  geom_bar(position = "dodge", color = "black") +
  labs(x = "Infected", y = "Count") +
  ggtitle("Bar Chart of infected and Treatment") +
  scale_fill_manual(values = c("drug" = "blue", "placebo" = "red")) +
  theme(legend.position = "bottom")

ggplot(Charlie, aes(x = treatment, fill = treatment)) +
  geom_bar(position = "dodge", color = "black") +
  labs(x = "Infected", y = "Count") +
  ggtitle("Bar Chart of Treatment") +
  scale_fill_manual(values = c("drug" = "blue", "placebo" = "red")) +
  theme(legend.position = "bottom")

```


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table2 <- xtabs(~infected + treatment, data=Charlie)
rowPerc(table2)
colPerc(table2)
```

People who ended up getting aids was about 25% of them. While 75% of the others took the placebo. That didn't get anything didn't take the drug and the other half took the placebo.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table2)
chisqtestGC(table2)
```

 Also Theres a big statistical difference that the P value of the drug did work. Theres large statistical evidence that this drug is effective in preventing aids for the group. Only 25% of the people got the drug while the other 75% did not. Most of the people who got aids took the placebo.



# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

